---
title: "ProcessRespirationData"
author: "Tim T. Rademacher"
date: "May 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Processing respiration data in R

This document describes how to process respiration data recorded with a LICOR-820 using a custom made application. The data is processed using functions from the `'RespChamberProc'` package.

## Load packages and dependencies

First, we install and load all dependencies (segmented and tibble) and the R functions from `'RespChamberProc'` package.
```{r installDependencies, include = FALSE}
install.packages ('segmented', repos = 'http://cran.us.r-project.org')
install.packages ('tibble',    repos = 'http://cran.us.r-project.org')
```
``` {load dependencies, echo = TRUE}
library ('segmented')
library ('tibble')
readFunctions <- function (x) { 
                   Rfiles <<- list.files (path = '../RespChamberProc/R/')
                   source (paste ('../RespChamberProc/R/', Rfiles [x], sep = ''))
}
res <- sapply (c(1:4,6:12), readFunctions)
```

## Read the data

In the working directory there is a folder `'data/'` with the data files. The files are in two formats. The first output format from the LICOR app, which can be read in using the `'readDat'` function form the `'RespChamberProc'` package. The second output format is from the `'CO2 Flux Application'` and can be read in using the `'readApp'` function. 

``` {r readDataFunction}
readApp <- function (fileName) {
  raw <- read.csv (file = paste ('../data/', fileName, sep = ''))
  names (raw) <- c ('Time.S', 'CO2', 'H20', 'Temperature', 'Pressure')
  return (raw)
}
```
``` {r setting, echo = F}
siteName   <- 'Exp2017'
chamberID  <- '02p4p2'
sampleDate <- as.POSIXct ('20180509', format = '%Y%m%d')
sampleTime <- as.POSIXct ('161645',   format = '%H%M%S')
```

The naming convention for the second type of files are `'G-'` for data (the respective metadata is file starting with `'M-'`), `'siteName'` such as `siteName`, `'chamberID'` such as `chamberID`, the `'sampleDate'` in `'%Y%m%d'` format such as `sampleDate`, and the `'sampleTime'` in `'%H%M%S'` format such as `sampleTime`. The examples would combine to "`r paste ('G-',siteName,'_',chamberID,'_',format (sampleDate, format = '%Y%m%d'),'_',format (sampleTime, format = '%H%M%S'),'.csv', sep = '')`", as these files are in csv format. 

``` {r readData}
fileName  <- paste ('G-',siteName,'_',chamberID,'_',format (sampleDate, format = '%Y%m%d'),'_',format (sampleTime, format = '%H%M%S'),'.csv', sep = '')
raw <- readApp (fileName)
```

The first processing step is to plot the CO2 data and determine the appropriate window to analyse.

``` {r plotRaw, echo = F}
plot (x = raw$Time.S,
      y = raw$CO2,
      xlab = 'time [s]',
      ylab = 'CO2 concentration [ppm]')
```

# Next step is to find a file with data that is not just a clean straight.